---
title: 'ggplot2: An Introduction'
author: "Jake Thompson"
date: "2/1/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## What is ggplot2?

ggplot2 is a data visualization package written by [Hadley Wickham](http://www.twitter.com/hadleywickham) that uses the "grammar of graphics." The grammar of graphics provides a consistent way to describe the components of graph, allowing us to move beyond specific types of plots (e.g., boxplot, scatterplot, etc.) to different elements that compose the plot. As the name would imply, the grammar of graphics is a language we can use to describe and build visualizations.

Today, we'll look at the basic syntax of ggplot2 graphics using simulated regression data.

```{r packages, message = FALSE}
library(ggplot2)
library(dplyr)
```


## The data

First, we'll define a function to generate data.

```{r sample_fn}
generate_data <- function(n, b0, b1, b2, bint, seed) {
  set.seed(seed)
  x1 <- rnorm(n = n, mean = 0, sd = 1)
  x2 <- sample(0:1, size = n, replace = TRUE)
  e <- rnorm(n = n, mean = 0, sd = sqrt(5))
  
  y <- b0 + (b1 * x1) + (b2 * x2) + (bint * x1 * x2) + e
  data_frame(y, x1, x2)
}
```

And then, we will use that function to generate a sample for our example.

```{r sample_data}
mlm_data <- generate_data(n = 1000, b0 = 3, b1 = 5, b2 = 3, bint = 4,
  seed = 9416)
```
